\section{Discussion}
\label{sec:discussion}

\subsection{Primary Findings}

The most robust empirical finding across all experiments is that structural market parameters dominate algorithmic design choices. In unconstrained settings (Experiments~1a--2b), the number of bidders yields the largest effect on revenue in every case. Table~\ref{tab:cross_exp_summary} ranks the top factors by absolute $t$-statistic for revenue across all six sub-experiments. The dominance of market thickness confirms the prediction of \citet{BulowKlemperer1996} that an additional bidder matters more than an optimal reserve, and extends it from static equilibria to settings where bidders are learning algorithms.

\begin{table}[H]
\centering
\caption{Cross-experiment summary of revenue drivers, format effects, and convergence. Factor rankings by absolute $t$-statistic for revenue; FPA and SPA columns show mean revenue in final evaluation period; convergence in rounds (episodes for Experiments~3a--3b).}
\label{tab:cross_exp_summary}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l l lll ccc c}
\toprule
& & \multicolumn{3}{c}{\textbf{Top Revenue Drivers ($|t|$)}}
& \multicolumn{3}{c}{\textbf{Revenue by Format}}
& \textbf{Conv.} \\
\cmidrule(lr){3-5} \cmidrule(lr){6-8} \cmidrule(lr){9-9}
\textbf{Exp} & \textbf{Algorithm}
& \textbf{Rank 1} & \textbf{Rank 2} & \textbf{Rank 3}
& \textbf{FPA} & \textbf{SPA} & \textbf{Gap (\%)}
& \textbf{Median} \\
\midrule
1a & Q-learn (const.) & \ExpOneARevFactorOne{} (\ExpOneARevFactorTOne) & \ExpOneARevFactorTwo{} (\ExpOneARevFactorTTwo) & \ExpOneARevFactorThree{} (\ExpOneARevFactorTThree) & \ExpOneARevMeanFPA & \ExpOneARevMeanSPA & $\ExpOneARevGapPct$ & \ExpOneAConvMedian \\
1b & Q-learn (affil.) & \ExpOneBRevFactorOne{} (\ExpOneBRevFactorTOne) & \ExpOneBRevFactorTwo{} (\ExpOneBRevFactorTTwo) & \ExpOneBRevFactorThree{} (\ExpOneBRevFactorTThree) & \ExpOneBRevMeanFPA & \ExpOneBRevMeanSPA & $\ExpOneBRevGapPct$ & \ExpOneBConvMedian \\
2a & LinUCB & \ExpTwoARevFactorOne{} (\ExpTwoARevFactorTOne) & \ExpTwoARevFactorTwo{} (\ExpTwoARevFactorTTwo) & \ExpTwoARevFactorThree{} (\ExpTwoARevFactorTThree) & \ExpTwoARevMeanFPA & \ExpTwoARevMeanSPA & $\ExpTwoARevGapPct$ & \ExpTwoAConvMedian \\
2b & Thompson & \ExpTwoBRevFactorOne{} (\ExpTwoBRevFactorTOne) & \ExpTwoBRevFactorTwo{} (\ExpTwoBRevFactorTTwo) & \ExpTwoBRevFactorThree{} (\ExpTwoBRevFactorTThree) & \ExpTwoBRevMeanFPA & \ExpTwoBRevMeanSPA & $\ExpTwoBRevGapPct$ & \ExpTwoBConvMedian \\
3a & Pacing (Mult.) & \ExpThreeARevFactorOne{} (\ExpThreeARevFactorTOne) & \ExpThreeARevFactorTwo{} (\ExpThreeARevFactorTTwo) & \ExpThreeARevFactorThree{} (\ExpThreeARevFactorTThree) & \ExpThreeARevMeanFPA & \ExpThreeARevMeanSPA & $\ExpThreeARevGapPct$ & \ExpThreeAConvMedian \\
3b & Pacing (PI) & \ExpThreeBRevFactorOne{} (\ExpThreeBRevFactorTOne) & \ExpThreeBRevFactorTwo{} (\ExpThreeBRevFactorTTwo) & \ExpThreeBRevFactorThree{} (\ExpThreeBRevFactorTThree) & \ExpThreeBRevMeanFPA & \ExpThreeBRevMeanSPA & $\ExpThreeBRevGapPct$ & \ExpThreeBConvMedian \\
\bottomrule
\end{tabular}}
\end{table}

Under budget constraints (Experiments~3a--3b), the budget multiplier takes over as the primary revenue driver. The structural shift from market thickness to budget tightness is the defining feature of constraint-driven environments.

The effect of auction format on revenue reverses between the two species of bid suppression. Under learning-driven suppression (Experiments~1a--2b), second-price auctions weakly dominate; under constraint-driven suppression (Experiments~3a--3b), first-price auctions dominate (Table~\ref{tab:cross_exp_summary}). This reversal qualifies the universal first-price penalty found by \citet{BanchioSkrzypacz2022}, who studied a single algorithm class. A platform choosing between formats must first assess which species of bid suppression its bidding agents are likely to produce.

The auction format $\times$ number of bidders interaction is significant in Experiments~1a and~2a but with opposite signs. Under Q-learning (Experiment~1a), additional bidders reduce the format effect; under LinUCB (Experiment~2a), additional bidders widen it. The measurement horizon also matters; when lifetime revenue is used instead of end-state revenue, the two formats become statistically indistinguishable in Experiment~1b.\footnote{First-price learning-phase premium: \ExpOneBFPAPremium; second-price: \ExpOneBSPAPremium. Lifetime revenue model: $|t| = \ExpOneBAllRevAuctionAbsT$, $p \ExpOneBAllRevAuctionPFmt$.} All primary findings survive comprehensive robustness checks (Appendix~\ref{sec:appendix_robustness}).

Quantile regression reveals that these mean effects mask distributional heterogeneity. In Experiment~1a, the auction format coefficient reverses sign across quantiles, indicating that first-price auctions compress the revenue distribution even though the average effect is negligible. In Experiments~2b, 3a, and~3b, the format effect grows toward the upper tail, so that the best-performing configurations are most sensitive to format choice (Appendix~\ref{sec:appendix_robustness}).

\subsection{Algorithm-Class Dependence of Secondary Effects}

Beyond the headline findings, a pervasive meta-pattern emerges across experiments. The direction and significance of nearly every secondary effect depends on the algorithm class. The affiliation parameter $\eta$ (spanning independent private to near-common values) shows no significant effect on revenue in Experiment~1b ($p = 0.056$). It becomes significant in both contextual bandit experiments (Experiment~2a, $t = -5.7$, $p < 0.001$; Experiment~2b, $t = -3.2$, $p = 0.002$). \citet{MilgromWeber1982} predict that second-price auctions should weakly dominate first-price auctions under affiliation; this prediction emerges only under contextual bandits, not under Q-learning. Reserve prices show similarly algorithm-dependent effects, significant and positive under Q-learning with constant valuations (Experiment~1a, $t = 5.1$) but significant and negative under Thompson Sampling (Experiment~2b, $t = -7.9$). Classical theory predictions are gated by the learning algorithm.

Efficiency outcomes under budget-constrained pacing illustrate a further dimension of algorithm-class dependence. Under utility-maximizing objectives, dual pacing agents achieve near-perfect allocative efficiency; under value-maximizing objectives with four bidders, efficiency drops substantially, with bid-to-value ratios exceeding 1.5.\footnote{Utility-maximizing efficiency ranges from \ExpThreeAUtilEffLow{} to \ExpThreeAUtilEffHigh{} across cells; value-maximizing four-bidder efficiency drops to \ExpThreeAValFourEffLow--\ExpThreeAValFourEffHigh.} All observed Price of Anarchy values fall well within the theoretical worst-case bound of~2 \citep{Aggarwal2019, Gaitonde2023} for both auction formats. However, the effect of auction format on the Price of Anarchy differs sharply between the two pacing mechanisms, with negligible contribution under dual pacing ($S_T = 0.005$) and dominant contribution under PI control ($S_T = 0.19$). Even within the same species of bid suppression, the mechanism choice matters for welfare.

Convergence speed varies substantially across algorithm classes. Table~\ref{tab:cross_exp_summary} compares convergence times across all experiments. In unconstrained settings, contextual bandits converge substantially faster than Q-learning, yet faster convergence does not translate into higher revenue. Experiment~2a converges far faster than Experiment~1a (\ExpTwoAConvMedian{} vs.\ \ExpOneAConvMedian{} rounds) but produces lower revenues.\footnote{Pacing experiments use a different convergence concept (within-episode budget depletion rather than cross-episode strategy settling), so direct comparison of convergence times between pacing and learning experiments is not meaningful.} First-price auctions produce higher winner entropy than second-price auctions under LinUCB (Experiment~2a), but this pattern does not replicate under Thompson Sampling (Experiment~2b) or Q-learning (Experiment~1b), where the auction format effect on winner entropy is not statistically significant. Higher exploration intensity reduces revenue under both bandit algorithms.

One finding is consistent across all algorithm classes: algorithmic tuning parameters contribute negligibly to revenue. Learning rate, discount factor, decay type, regularisation, memory decay, and aggressiveness each achieve $S_T < 0.04$ in every experiment where they are varied. Market structure dominates algorithm configuration universally.

\subsection{Implications}

The most effective way to improve seller revenue is ensuring sufficient competition. In unconstrained settings, increasing the number of bidders improves revenue more than any other intervention tested, including switching auction formats. In budget-constrained environments, relaxing budget tightness has the largest impact. The effect of auction format is context-dependent; sellers should choose format based on the prevailing bidding technology rather than adopting a blanket preference.

The dominance of structural market parameters over algorithmic design choices carries a direct implication for competition policy. Structural market interventions, such as lowering barriers to entry in auction platforms or preventing excessive market concentration among bidders, are likely to be more effective than algorithm-specific regulation. Across all unconstrained experiments, adding bidders produces a larger revenue improvement than any algorithmic intervention, and under budget constraints, relaxing budget tightness dominates all other factors. The context-dependence of auction format effects complicates prescriptive regulation. A blanket mandate requiring a particular auction format ignores the finding that the welfare consequences of format choice depend on the learning algorithm, the value environment, and the presence of budget constraints. Under Q-learning with constant valuations, the two formats produce statistically indistinguishable revenue. Under contextual bandits with affiliated values, second-price auctions yield higher revenue. Under budget-constrained pacing, first-price auctions perform better. This heterogeneity supports outcome-based monitoring over rule-based mandates \citep{OECD2023, Johnson2023}. The ranked-effects tables produced by the factorial analysis quantify the marginal contribution of each factor, offering a transparent and reproducible evidentiary basis that could support such outcome-based regulatory approaches \citep{Harrington2018, Gal2019}.

\subsection{Practitioner's Guide}
\label{sec:practitioners_guide}

The factorial laboratory is a general-purpose diagnostic for algorithmic markets, not a method tied to the specific algorithms or auction formats studied here. Any researcher who suspects that autonomous bidding agents may suppress prices, inflate costs, or distort allocations can apply the same workflow. The first step is to characterise the algorithm-mechanism pair under study. Nearly every secondary finding depends on the algorithm class, so the pair must be identified before any design decisions are made. The researcher should determine whether agents are learning values from scratch (Q-learning, contextual bandits), managing budget constraints across repeated auctions (multiplicative pacing, PI controllers), or following a different approach altogether (deep reinforcement learning, transformer-based bidders, LLM agents). The market rules matter equally. Sealed-bid versus open formats, first-price versus second-price payment rules, reserve price policies, and single-item versus combinatorial allocation all shape which theoretical predictions serve as benchmarks and which factors are worth screening.

The second step is to survey the relevant theory and empirical literature for candidate factors, plausible ranges, and theoretical predictions. Classical auction theory supplies equilibrium benchmarks against which simulated outcomes can be compared, including revenue equivalence under independent private values, revenue rankings under affiliation, and the marginal value of an additional bidder. The algorithmic collusion literature identifies mechanism-specific factors such as exploration parameters, discount rates, synchronisation modes, and state representations. The autobidding literature highlights budget multipliers, pacing objectives, and the distinction between value-maximising and utility-maximising agents. Each candidate factor should have a theoretical or empirical basis for inclusion and a range spanning practically relevant levels. Section~\ref{sec:literature} discusses the four bodies of literature this paper draws on.

The third step is to screen the candidate factors down to a manageable set. No single experiment can test everything, so the screening stage selects the factors most likely to matter. A sequential approach begins with a large screening design that spans many factors at two levels each, then follows up on active factors with focused designs that add levels or new response variables. The key discipline is to separate structural market parameters (number of bidders, reserve prices, value environments) from algorithmic tuning parameters (learning rates, discount factors, exploration intensity), because the two categories may operate at different effect scales. Factor levels should span the range encountered in practice rather than being chosen for theoretical convenience. Table~\ref{tab:params} lists the factor definitions and ranges used in this paper.

The fourth step is to select an experimental design that provides orthogonal or near-orthogonal estimation of main effects and interactions. Two-level factorial designs, whether full or fractional, are the standard choice for screening. When some factors have more than two natural levels, as with the affiliation parameter spanning independent, moderate, and common values, mixed-level designs accommodate the extra levels while preserving balance. Space-filling designs such as Latin hypercubes or Sobol sequences are an alternative when factors are continuous and the researcher wants to map the response surface rather than screen for active factors. The choice depends on the research question; factor identification favours factorials, while functional-form mapping favours space-filling designs. Replication should scale inversely with design size to supply pure-error degrees of freedom for lack-of-fit testing. Section~\ref{sec:inference} details the designs used in this paper.

Estimation combines ANOVA with a surrogate model. An OLS model with main effects and all two-way interactions, fit using effects coding ($-1$/$+1$), guarantees orthogonal estimation of all estimable effects (Section~\ref{sec:appendix_inference}). The linear model should always be accompanied by a nonparametric surrogate, such as gradient-boosted trees with cross-validation, to benchmark model adequacy; a meaningful $R^2$ gap signals detectable nonlinearity that the linear model misses. Pareto charts of absolute $t$-statistics and half-normal probability plots are the primary tools for separating active effects from noise, providing a visual complement to the numerical significance tests.

The final step is to report results in a structured way. Variance decomposition through Sobol indices, both first-order $S_i$ and total-order $S_{T_i}$, quantifies each factor's contribution to outcome variability; for balanced factorials these coincide with the ANOVA sum-of-squares decomposition (Section~\ref{sec:appendix_sensitivity}). Effect direction and magnitude come from the OLS coefficients. Multiple-testing corrections are essential when many effects are tested simultaneously; family-wise error rate control via Holm--Bonferroni and false discovery rate control via Benjamini--Hochberg guard against false positives. Quantile regressions at multiple percentiles reveal whether effects are uniform across the outcome distribution or concentrated in the tails. Power analysis establishes what the design can and cannot detect. Together, these diagnostics provide a transparent, reproducible evidentiary basis for ranking the factors that matter most. Appendix~\ref{sec:appendix_robustness} presents the full battery applied in this paper. Figure~\ref{fig:lab_pipeline} summarises this workflow.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  node distance=0.45cm,
  block/.style={rectangle, draw, rounded corners, minimum width=2.0cm,
                minimum height=0.9cm, text centered, font=\small},
  arrow/.style={-{Stealth[length=2mm]}, thick},
]
  % Nodes
  \node[block] (spec) {Specify};
  \node[block, right=of spec] (design) {Design};
  \node[block, right=of design] (sim) {Simulate};
  \node[block, right=of sim] (est) {Estimate};
  \node[block, right=of est] (val) {Validate};
  \node[block, right=of val] (dec) {Decide};

  % Forward arrows
  \draw[arrow] (spec) -- (design);
  \draw[arrow] (design) -- (sim);
  \draw[arrow] (sim) -- (est);
  \draw[arrow] (est) -- (val);
  \draw[arrow] (val) -- (dec);

  % Annotations below
  \node[below=0.15cm of spec, font=\scriptsize, align=center] {algorithm,\\mechanism, factors};
  \node[below=0.15cm of design, font=\scriptsize, align=center] {factorial or\\space-filling};
  \node[below=0.15cm of sim, font=\scriptsize, align=center] {run auction\\experiments};
  \node[below=0.15cm of est, font=\scriptsize, align=center] {OLS + surrogate\\benchmarking};
  \node[below=0.15cm of val, font=\scriptsize, align=center] {robustness\\checks};
  \node[below=0.15cm of dec, font=\scriptsize, align=center] {Sobol indices,\\policy guidance};

  % Feedback arrow from Validate back to Specify
  \draw[arrow, rounded corners=10pt, gray]
    (val.south) -- ++(0, -1.3) -| (spec.south)
    node[pos=0.25, below, font=\scriptsize] {iterate};
\end{tikzpicture}
\caption{Laboratory pipeline for diagnosing algorithmic markets. The six stages correspond to the workflow described in the Practitioner's Guide. A feedback loop from validation to specification supports iterative refinement of factor sets and design choices.}
\label{fig:lab_pipeline}
\end{figure}

\subsection{Limitations and Future Directions}

The qualitative patterns documented here are likely to hold in settings characterised by repeated interactions, autonomous learning, and sealed-bid formats, conditions that closely approximate programmatic advertising markets. External validity to specific real-world markets requires empirical validation with proprietary auction data. The algorithms studied here are stylised; the experiments do not include deep reinforcement learning, LLM-based bidders, or transformer architectures. Agents have no explicit communication channels, episode lengths are fixed, bidder populations are symmetric, and no combinatorial formats are tested. Settings involving few-shot auctions or human bidders may produce qualitatively different outcomes.

This work can be extended in three directions. Response surface designs can map the precise functional forms of the factors identified as important by the screening factorials, moving from factor identification to functional-form estimation. Multi-agent architectures beyond tabular methods and PI controllers, including transformers, LLM-based bidders, and deep reinforcement learning, would test whether the dominance of structural parameters persists when algorithmic complexity increases. Empirical validation with proprietary auction data would test whether the laboratory patterns replicate in real advertising exchanges.
