\section{Experimental Design and Statistical Methods}
\label{sec:experiments}

We design four experiments (with Experiment~III split into two sub-experiments by algorithm class) to study repeated sealed-bid auctions under varying valuations and learning strategies. All factors are coded as $-1$ (low level) and $+1$ (high level) for statistical analysis; this effects coding ensures orthogonal estimation of main effects and interactions.

In all experiments, $n$ bidders submit discrete bids $b_i\in[0,1]$ each round. A reserve price $r\ge 0$ may invalidate bids below $r$; if no valid bid meets $r$, no sale occurs and revenue is zero. The highest valid bid wins.\footnote{When multiple bids tie for the highest valid bid, a single winner is selected uniformly at random among those tied.} We compare first-price (winner pays her own bid) and second-price (winner pays the second-highest bid) formats. Each experiment allows multiple parameter configurations and collects the outcome metrics listed below.

\subsection{Experiment Descriptions}

Experiment~I deploys Q-learning agents with constant valuations ($v_i = 1$) in a $2^{10-1}$ Resolution~V half-fraction spanning 10 factors (learning rates, discount factors, exploration modes, update synchronisation, number of bidders, reserve prices) and 1{,}024 observations. Under independent private values, revenue equivalence \citep{Vickrey1961} predicts that first- and second-price auctions yield identical expected revenue. The experiment tests whether Q-learning preserves this equivalence or breaks it.

Experiment~II generalises Q-learning to affiliated valuations via $v_i = (1 - 0.5\eta)s_i + 0.5\eta \frac{1}{n-1}\sum_{j \neq i} s_j$, where $\eta \in \{0, 0.5, 1\}$ spans private to near-common values. The revenue ranking theorem \citep{MilgromWeber1982} predicts SPA $\geq$ FPA under affiliation. The $3 \times 2^3$ mixed-level design crosses auction format, affiliation, number of bidders, and state information, with Q-learning hyperparameters fixed at levels identified in Experiment~1, testing whether Q-learning preserves this revenue ordering.

Experiment~IIIa replaces Q-learning with LinUCB contextual bandits under the same affiliated valuation model. LinUCB uses optimism-based exploration, while Thompson Sampling uses posterior sampling, creating a natural contrast for how exploration style affects outcomes. Bandit algorithms should exploit contextual structure better than tabular Q-learning; the experiment tests whether this improves seller outcomes.

The $3 \times 2^7$ mixed-level design for IIIa has 9 factors (7 binary plus $\eta$) and 384 cells, replicated twice for 768 observations. It includes regularisation ($\lambda$) and memory decay as factors, both of which strongly affect LinUCB performance. Experiment~IIIb deploys Contextual Thompson Sampling with a smaller $3 \times 2^5$ design (7 factors, 96 cells, 192 observations). It drops regularisation and memory decay because the Bayesian posterior update renders the prior negligible after approximately 20 rounds.\footnote{Thompson Sampling explores via posterior sampling, which is inherently stochastic even as the posterior tightens.} Separating the two algorithms avoids wasting design cells on factors that are structurally irrelevant to one algorithm class.

Experiment~IVa introduces budget-constrained autobidding agents using multiplicative dual pacing, implementing the framework of \citet{Balseiro2019Learning}. Each of $n \in \{2, 4\}$ advertisers participates in 100 episodes of 1{,}000 auctions each, with budgets regenerating between episodes while dual variables persist. The theoretical worst-case Price of Anarchy is bounded by~2 for both auction formats \citep{Aggarwal2019, Gaitonde2023}. The $2^6$ full factorial crosses auction format, bidder objective (value- vs.\ utility-maximiser), number of bidders, budget multiplier (tight vs.\ loose), reserve price, and value dispersion, replicated across 8 seeds for 512 observations.

Experiment~IVb retains the budget-constrained environment of Experiment~IVa but replaces multiplicative dual pacing with a proportional-integral (PI) controller. The PI agent bids $\lambda \cdot v$ and updates $\lambda$ additively based on cumulative spending error. The experiment tests whether the control law matters or whether budget structure is the primary driver of outcomes. The $2^6$ full factorial crosses auction format, controller aggressiveness (conservative vs.\ aggressive PI gains), number of bidders, budget multiplier, reserve price, and value dispersion, replicated across 8 seeds for 512 observations. Detailed parameter tables for each experiment appear in Appendix~\ref{sec:appendix_design}. Table~\ref{tab:params} summarises the parameters used across all experiments.

\subsection{Design Rationale and Alternatives}

All experiments use two-level factorial designs (with one three-level factor for the affiliation parameter $\eta$ in Experiments~2--3b). This choice reflects a screening objective: identifying which factors matter and ranking their relative importance, rather than mapping precise functional forms. Two-level designs are the most efficient screening tool, requiring only $2^k$ or $2^{k-p}$ runs to achieve full orthogonality among main effects and low-order interactions \citep{Box2005, Montgomery2017}. With 6 to 10 factors per experiment, the two-level factorial provides the maximum information per simulation run about which factors drive outcomes.

At the economical end of the design spectrum, fractional factorials and Plackett-Burman designs minimise run counts at the cost of aliasing higher-order interactions with main effects. Our Resolution~V half-fraction in Experiment~1 aliases only four-factor and higher interactions, preserving clean estimation of all main effects and two-factor interactions. Definitive screening designs \citep{JonesNachtsheim2011} offer an even leaner alternative, estimating main effects, quadratic effects, and some two-factor interactions in only $2k+1$ runs, making them attractive when simulation budgets are tight and curvature detection is desired without a full response surface experiment.

When the goal shifts from screening to optimisation, richer designs become appropriate. Central composite designs \citep{BoxWilson1951} and Box-Behnken designs \citep{BoxBehnken1960} add intermediate factor levels to fit second-order response surfaces, enabling identification of optima within the factor space. For computer experiments where the response surface may be highly nonlinear, space-filling designs such as Latin hypercube sampling \citep{McKayBeckmanConover1979} and maximin distance designs \citep{SantnerWilliamsNotz2018} spread evaluation points across the entire factor space, providing the uniform coverage needed for Gaussian process emulation or other nonparametric metamodels.

Our experimental programme follows \citeauthor{Box2005}'s sequential experimentation philosophy: screen first with efficient designs, then investigate further with richer designs on the factors that matter. Experiments~1 and~3a screen many factors with large factorial designs; their findings inform the reduced factor sets in Experiments~4a and~4b. A natural extension of this work would apply response surface or space-filling designs to the dominant factors identified here, mapping the precise functional relationship between, for example, the number of bidders or the budget multiplier and auction revenue.

\subsection{Parameter Ranges}

\begin{table}[H]
\centering
\caption{Parameter Ranges and Their Usage Across Experiments}
\label{tab:params}
\small
\begin{tabular}{l l c c c c c c}
\toprule
\textbf{Name} & \textbf{Description} & \textbf{E1} & \textbf{E2} & \textbf{E3a} & \textbf{E3b} & \textbf{E4a} & \textbf{E4b}\\
\midrule
$\alpha$ & Q-learning rate & \checkmark & & & & & \\
$\gamma$ & Discount factor & \checkmark & & & & & \\
$\varepsilon$ & E-greedy exploration prob & \checkmark & & & & & \\
Boltzmann & Softmax exploration & \checkmark & & & & & \\
$c$ / $\sigma^2$ & Bandit exploration param &  &  & \checkmark & \checkmark & & \\
$\lambda$ & Regularisation &  &  & \checkmark & & & \\
\textit{Memory decay} & Observation weighting &  &  & \checkmark & & & \\
$r$ & Reserve price & \checkmark & & \checkmark & \checkmark & \checkmark & \checkmark \\
$n$ & Number of bidders & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
$\eta$ & Affiliation parameter &  & \checkmark & \checkmark & \checkmark & & \\
\textit{Episodes} & Total training rounds & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
\textit{Sync/Async} & Q-learning modes & \checkmark & & & & & \\
\textit{State info} & State features & \checkmark & \checkmark & & & & \\
\textit{Context richness} & Context features &  &  & \checkmark & \checkmark & & \\
\textit{Decay type}  & Epsilon decay schedule         & \checkmark & & & & & \\
\textit{Objective}   & Value- vs.\ utility-maximiser  & & & & & \checkmark & \\
\textit{Aggressiveness} & PI controller gain scaling & & & & & & \checkmark \\
$m$ & Budget multiplier & & & & & \checkmark & \checkmark \\
$\sigma$ & Value dispersion & & & & & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Outcome Metrics}

Our core outcome metrics are consistent across all experiments. We measure the \emph{average revenue in later rounds}, which reflects how well the auction performs after strategies have stabilised.\footnote{Specifically, we average revenue over the final 1{,}000 episodes for Experiments~1--3b and over all post-burn-in episodes for Experiments~4a--4b.} We also determine the \emph{time to converge} as the earliest round after which revenue remains near its final level.\footnote{Defined as the earliest round after which the rolling-average revenue stays within $\pm 5\%$ of its final mean.} We compute a \emph{no-sale rate} by noting the fraction of rounds in which no valid bid exceeds the reserve. We track \emph{price volatility} by taking the sample standard deviation of winning bids in later rounds. Finally, we measure \emph{winner entropy} to assess whether outcomes concentrate among few bidders, calculating the Shannon entropy of the empirical distribution of winners. Table~\ref{tab:outcomes} provides a concise summary of these common metrics that allow systematic comparisons of convergence speed, bidding stability, and revenue performance across all experiments.

\begin{table}[h]
\centering
\caption{Outcome metrics used in all experiments.}
\label{tab:outcomes}
\begin{tabular}{l l}
\toprule
\textbf{Metric} & \textbf{Description}\\
\midrule
Average revenue (later rounds) & Mean revenue in the final 1000 rounds \\
Time to converge & Round at which revenue stays in a $\pm 5\%$ band \\
No-sale rate & Fraction of rounds with all bids below $r$\\
Price volatility & Standard deviation of winning bids in later rounds \\
Winner entropy & Shannon entropy of bidder identity distribution \\
\bottomrule
\end{tabular}
\end{table}

Experiments~2, 3a, 3b, 4a, and~4b introduce additional response variables specific to their designs; these are defined in Appendix~\ref{sec:appendix_summary} and Appendix~\ref{sec:exp4_appendix}, respectively.

\subsection{Factorial Design Framework}

The factorial designs described above are analysed using ordinary least squares with effects coding ($x_i \in \{-1, +1\}$). This coding ensures orthogonality: each main effect and two-way interaction can be estimated independently of all others, enabling clean attribution of outcome variation to specific factors. Experiment~1's $2^{10-1}$ half-fraction has Resolution~V, ensuring that all main effects and two-way interactions are estimable without bias from three-way terms, which are assumed negligible under effect sparsity.\footnote{In the smaller mixed-level designs (Experiments~2--3b), the degrees of freedom for error scale directly with replication count. Experiment~2's $3 \times 2^3$ design has 24 cells and 16 model parameters; with eight replicates the residual degrees of freedom are adequate. HC3 robust standard errors and wild bootstrap $p$-values (Appendix~\ref{sec:appendix_robustness}) confirm that findings are stable across all sub-experiments.} Replication in all experiments enables pure error estimation and lack-of-fit testing.

The stochastic variation in our experiments arises from seed-induced randomness in exploration, valuation draws, and tie-breaking. Inference is over the distribution of these random environments, not over a population of real markets. This is methodologically standard in computational economics and agent-based modeling \citep{Tesfatsion2006}. Factorial designs applied to stochastic simulations yield unbiased effect estimates under the same assumptions as physical experiments, with seed variation playing the role of experimental noise.

\subsection{Estimation}

For each response variable $Y$ (e.g., average revenue, convergence time, price volatility), we fit a linear model with main effects and all two-way interactions via ordinary least squares (OLS):
\begin{align}
  Y_i = \beta_0 + \sum_{j=1}^{k} \beta_j\, x_{ji} + \sum_{1 \le j < l \le k} \beta_{jl}\, x_{ji}\, x_{li} + \varepsilon_i,
  \label{eq:ols}
\end{align}
where $x_{ji} \in \{-1, +1\}$ is the coded level of factor $j$ in observation $i$, $\beta_j$ represents the main effect of factor $j$, and $\beta_{jl}$ captures the two-way interaction between factors $j$ and $l$. Under the coded parameterisation, $\beta_j$ equals half the change in mean response when factor $j$ moves from its low to its high level, holding all other factors at their centre values. We use Type~III ANOVA decomposition to assess the marginal contribution of each term,\footnote{Under the $\pm 1$ effects coding in a balanced factorial, the design matrix is orthogonal ($\mathbf{X}^\top\mathbf{X}$ is diagonal), so the OLS $t$-test for each coefficient is algebraically equivalent to the Type~III ANOVA $F$-test for the same term ($F = t^2$), yielding identical $p$-values. This equivalence depends on orthogonality; under $0/1$ dummy coding the columns would be correlated, main effect estimates would be conditional on the reference level rather than marginal, and the two tests would diverge.} testing each coefficient against the null $H_0\!: \beta = 0$ via the $t$-statistic $\hat{\beta}/\text{SE}(\hat{\beta})$.\footnote{The ANOVA $F$-tests are unadjusted for multiplicity. In the factorial design tradition \citep{Box2005}, the primary screening tools are the half-normal probability plot and Lenth's pseudo-standard-error method \citep{Lenth1989}, both of which identify active effects without requiring an independent error estimate. Formal multiplicity corrections (Holm--Bonferroni, Benjamini--Hochberg) are applied in the robustness analysis described below.} Model adequacy is validated by comparing OLS $R^2$ to a gradient-boosted machine (LightGBM) $R^2$ trained with five-fold cross-validation: gaps below 0.05 support adequate linear approximation; gaps above 0.05 reveal detectable nonlinearity that warrants checking whether effect rankings change under the nonparametric model.

\subsection{Robustness and Model Adequacy}

We validate all findings through a comprehensive suite of robustness checks. For standard errors and multiple testing, we compute HC3 heteroskedasticity-consistent standard errors \citep{MacKinnonWhite1985}, Holm--Bonferroni and Benjamini--Hochberg corrections, and Rademacher wild bootstrap $p$-values for the top effects.

For specification and model adequacy, we run quantile regressions at five percentiles, lack-of-fit $F$-tests with pure error from replication, leave-one-out cross-validated $R^2$ via the PRESS statistic, LightGBM nonparametric benchmarking, and LASSO variable selection.

No multiplicity correction is applied across the six sub-experiments. The probability that the number of bidders ranks first among $k$ factors in all six independent sub-experiments by chance is bounded above by $(1/k)^6$, which for $k \geq 3$ is below 0.0014. This permutation argument provides conservative evidence that the ordinal ranking is unlikely to arise by chance.\footnote{$t$-statistic magnitudes are not comparable across experiments due to differences in sample size, model specification, and error variance; the comparison is strictly ordinal.}

We compute minimum detectable effects (MDE) at 80\% power and $\alpha = 0.05$:
\begin{align}
  \text{MDE}_{j} = \bigl(t_{\alpha/2,\,\nu} + z_{0.80}\bigr) \times \frac{\hat{\sigma}}{\sqrt{\sum_{i} x_{ij}^{2}}},
  \label{eq:mde}
\end{align}
where $\hat{\sigma}$ is the residual standard deviation, $\nu$ is the residual degrees of freedom, and $\sum_{i} x_{ij}^{2}$ is the sum of squared coded values for factor~$j$. For binary $\pm 1$ factors this reduces to $\hat{\sigma}/\sqrt{n}$; for orthogonal polynomial contrasts the sum of squares differs by column (Appendix~\ref{sec:appendix_inference}). MDEs range from 2--5\% of the mean response, indicating that the designs have sufficient power to detect practically meaningful effects. Full methodological details appear in Appendix~\ref{sec:appendix_inference}, and per-experiment diagnostics are reported in Appendix~\ref{sec:appendix_robustness}.
