\section{Related Work}
\label{sec:literature}

This paper sits at the intersection of auction theory, algorithmic collusion, autobidding, and competition policy. The theoretical predictions of rational auction models supply the benchmarks against which algorithmic behaviour is measured; the collusion literature documents the mechanisms by which learning agents depart from those benchmarks; the autobidding literature asks how budget-constrained pacing reshapes welfare and efficiency; and legal scholarship clarifies why autonomous algorithmic convergence falls outside current enforcement frameworks. The remainder of this section reviews each body of work in turn and identifies the gaps that motivate this paper.

\subsection{Auction Theory and Equilibrium Predictions}

The classical auction-theoretic results serve as null hypotheses throughout Experiments~1--3. \citet{Vickrey1961} establishes revenue equivalence under independent private values: sealed-bid first-price and second-price auctions yield identical expected revenue when bidders are symmetric and risk-neutral. \citet{MilgromWeber1982} extend the analysis to affiliated signals, showing that the second-price auction generates weakly higher expected revenue than the first-price auction because the linkage principle mitigates the winner's curse. \citet{BulowKlemperer1996} demonstrate that an additional bidder raises expected revenue more than any optimal reserve price, a finding that motivates the inclusion of market thickness as a factor in every experiment. \citet{Krishna2009} provides the textbook synthesis on which the equilibrium benchmarks in this paper are based.

In repeated settings, these one-shot predictions can break down. The Folk Theorem implies that if bidders are sufficiently patient, any feasible and individually rational payoff vector, including symmetrically suppressed bids or asymmetric bid rotation, can be sustained as a subgame-perfect equilibrium. \citet{Ivaldi2003} characterise the determinants of tacit collusion in repeated oligopoly games: collusion is facilitated by higher discount rates, fewer participants, symmetric market conditions, high-frequency interaction, and greater transparency. Algorithms amplify the last three factors, interacting at millisecond frequency, observing auction outcomes in full, and processing information with perfect recall, which explains much of the concern about algorithmic collusion in auction markets.

\subsection{Algorithmic Collusion}

A growing literature studies whether reinforcement-learning agents converge to supra-competitive outcomes, primarily in pricing games. \citet{Calvano2020} show that Q-learning agents in simultaneous Bertrand competition develop reward-punishment strategies resembling those predicted by the Folk Theorem, sustaining prices above the Nash equilibrium. \citet{Klein2021} extends this to sequential pricing, finding that the granularity of the action space matters: finer grids facilitate collusion by enabling small, retaliatory price adjustments. \citet{Hansen2021} demonstrate that collusion can arise from misspecified demand prediction and correlated price experimentation, without any strategic intent. \citet{Calvano2023} introduce a distinction between genuine collusion, sustained by learned punishment schemes, and spurious collusion, where agents converge to high prices simply because exploration is too limited to discover profitable deviations. \citet{Douglas2024} formalise this insight, proving analytically that deterministic bandit algorithms always converge to collusive outcomes in repeated games, while persistent stochastic exploration prevents such convergence even when agents observe only their own payoff history.

Fewer studies examine auctions specifically. \citet{BanchioSkrzypacz2022} compare first- and second-price auctions with Q-learning bidders and find that the first-price format is more prone to bid suppression. \citet{WaltmanKaymak2008} study reinforcement learning in Cournot-style and reverse auction settings, showing that mixed-strategy equilibria can be attained in simple cases. \citet{Bandyopadhyay2008} reach similar conclusions for reverse auctions. \citet{Tellidou2007} find that tacit collusion is sustained even under competitive conditions in simulated electricity markets. This paper extends \citet{BanchioSkrzypacz2022} by varying ten or more factors simultaneously within a factorial design, rather than comparing formats in a fixed environment.

Several papers focus on the role of exploration and algorithmic sophistication. \citet{Abada2022} and \citet{Dolgopolov2021} show that the choice of exploration strategy, whether epsilon-greedy, Boltzmann, or upper-confidence-bound, materially affects the likelihood of collusive convergence. \citet{Asker2022} highlight synchronous versus asynchronous updating as a further determinant. \citet{Hettich2021}, \citet{Han2022}, and \citet{Zhang2021} extend the analysis to deep reinforcement learning, comparing different network architectures and sampling strategies. A common criticism of this simulation-based literature is that the algorithms studied bear little resemblance to those deployed in real markets \citep{KuhnTadelis2017, Schwalbe2018}. Every paper cited above uses Q-learning, simple bandit algorithms, or shallow reinforcement-learning variants, yet the dominant bidding systems on major auction platforms employ budget pacing, dual-variable optimisation, and proportional-integral controllers that solve fundamentally different problems: allocating a finite budget across auctions rather than learning value estimates from repeated play. The distinction matters for policy. The headline finding that first-price auctions are more vulnerable to algorithmic bid suppression than second-price auctions \citep{BanchioSkrzypacz2022} was derived entirely from Q-learning experiments. If that conclusion is an artefact of the algorithm class rather than an intrinsic property of the auction format, then platform design recommendations built on it may be misleading. This paper addresses the critique directly by embedding Q-learning, contextual bandits, and budget-constrained pacing within a common factorial framework, so that the robustness of each finding to the choice of algorithm can be assessed rather than assumed.

Recent theoretical work has formalised the mechanisms underlying algorithmic collusion. \citet{BanchioMantegazza2022} identify ``spontaneous coupling'' as a key driver: when Q-learning agents explore infrequently, correlated estimation errors synchronise their play, creating stochastic cycles that sustain supra-competitive outcomes without deliberate reward-punishment schemes. \citet{Bertrand2025} strengthen this result by proving that Q-learners with one-step memory converge to the cooperative Pavlov (win-stay, lose-shift) policy in iterated prisoner's dilemmas, establishing the first formal convergence guarantee for collusion by standard stochastic Q-learning. \citet{Calzolari2021} extend the analysis to settings with imperfect monitoring of competitors' actions, showing that the degree of market observability interacts with algorithmic learning to determine whether collusive outcomes can be sustained. \citet{Zhang2025Noise} offers a distinct perspective by modelling collusion as a belief-averaging process mediated by shared market data, and demonstrates that calibrated noise injection can disrupt coordination without eliminating the informational benefits that pricing algorithms provide to individual firms.

The scope of algorithmic collusion research continues to expand beyond tabular reinforcement learning and beyond folk-theorem punishment schemes. \citet{Fish2024} show that large language models autonomously reach supra-competitive prices in oligopoly experiments, and that seemingly innocuous prompt modifications substantially alter collusive outcomes. \citet{Arunachaleswaran2025} demonstrate that supra-competitive prices can arise even when algorithms explicitly lack the capacity for threats: if one firm deploys any no-regret learning algorithm and the other merely optimises its own revenue, near-monopoly prices emerge without either party encoding punishments or responding to competitor actions. These findings collectively establish that supra-competitive convergence is a robust phenomenon, driven by fundamental properties of adaptive algorithms rather than by the idiosyncrasies of any particular implementation, and raise distinct challenges for detection because the strategies producing collusive outcomes need not exhibit the suspicious patterns that enforcement frameworks are designed to identify.

Empirical evidence motivates the concern but lacks controlled experimental decomposition. \citet{Chen2016} find that over a third of Amazon best-sellers use algorithmic pricing. \citet{Assad2020} document margin increases of 9--28\% following the adoption of algorithmic pricing in German gasoline markets. \citet{BrownMacKay2023} estimate 10\% price increases among large online retailers using pricing algorithms. These studies establish the real-world prevalence and magnitude of algorithmic effects but cannot isolate which algorithmic or market features drive the observed outcomes.

The experimental literature with human bidders provides a complementary reference point. \citet{KagelLevin2016} survey the evidence and note that bids are generally above Nash predictions in both first- and second-price auctions. \citet{Levin1996} show that even experienced bidders fall prey to the winner's curse when signals are noisy. \citet{Keser1996} and \citet{Neugebauer2007} document declining prices in sequential auctions of identical items. These departures from theory in human experiments make it all the more important to understand the distinct, and often larger, departures that arise when algorithms replace human bidders.

\subsection{Autobidding, Pacing, and Platform Welfare}

A parallel literature, originating in algorithmic game theory rather than industrial organisation, examines autobidding and pacing in online advertising auctions. Where the algorithmic collusion literature asks whether learning agents converge to supra-competitive prices, this body of work asks how much welfare is lost when budget-constrained advertisers delegate bidding to automated pacing agents. \citet{Aggarwal2019} show that uniform bid-scaling is the optimal single-agent strategy in truthful auctions and establish a worst-case Price of Anarchy of~2 for second-price auctions under budget constraints. \citet{Conitzer2022FPPE} prove that first-price pacing equilibria are unique and computable via the Eisenberg--Gale convex programme, connecting autobidding to classical competitive equilibrium theory. \citet{Deng2021TowardsEfficient} show that first-price auctions achieve a Price of Anarchy of~1 under return-on-spend constraints, though this result does not extend to budget-constrained settings where the applicable bound remains~2. \citet{Aggarwal2024Survey} provide a recent survey of this rapidly growing field.

On the algorithmic side, \citet{Balseiro2019Learning} develop dual-based pacing algorithms in which each agent adjusts a shadow price on its budget constraint, suppressing bids when overspending and raising them when underspending, and prove $O(\sqrt{T})$ individual regret guarantees. \citet{Gaitonde2023} strengthen this by showing that gradient-based pacing achieves at least half the optimal liquid welfare without requiring convergence to equilibrium, a result that holds for any core auction format. \citet{PaesLeme2024} demonstrate that autobidding dynamics can nonetheless exhibit bi-stability and periodic orbits even in simple market structures, raising the question of whether pacing systems reach equilibrium in practice. This literature characterises efficiency from the bidder's perspective but has not examined whether budget-mediated bid suppression produces systematic revenue shortfalls for the seller, nor whether the auction-format and market-thickness interactions observed in learning-driven settings persist when bid suppression arises from budget rationing rather than Q-value underestimation.

Recent work has established format-specific welfare guarantees that hold even without convergence to equilibrium. \citet{FikiorisTardos2023} prove that in sequential first-price auctions with budget-constrained no-regret learners, liquid welfare is within a factor of approximately~2.41 of the optimum, whereas in second-price auctions the ratio can be arbitrarily bad even when all bidders satisfy no-regret guarantees. This sharp format comparison provides a welfare-theoretic rationale for the ongoing industry transition from second-price to first-price auction formats. \citet{Lucier2023PacingDynamics} extend the welfare analysis to the practically relevant setting with joint budget and return-on-investment constraints, proposing a gradient-based autobidding algorithm that achieves at least half the optimal liquid welfare for any core auction format without requiring convergence to equilibrium. These results strengthen the theoretical motivation for Experiments~4a and~4b, which test whether pacing algorithms realise these welfare guarantees in practice across varied market configurations. The theoretical predictions also create a direct tension with the algorithmic collusion literature. \citet{FikiorisTardos2023} prove that first-price auctions are strictly superior for liquid welfare under budget-constrained pacing, and \citet{Conitzer2022FPPE} show that first-price pacing equilibria are unique and efficient, whereas \citet{BanchioSkrzypacz2022} find that first-price auctions are more vulnerable to bid suppression under Q-learning. Whether this apparent contradiction reflects a genuine difference in auction-format effects across algorithm classes, or merely an artefact of the stylised settings in which each result was obtained, is an open empirical question that the present paper is designed to resolve.

\subsection{Legal and Regulatory Context}

\citet{EzrachiStucke2017} distinguish four modes of algorithmic participation in collusion, ranging from algorithms as messengers carrying out human-devised cartels to fully autonomous learning agents that converge to collusive outcomes without agreement or intent. \citet{Mehra2016} frames the fundamental challenge as an ``agreement gap'': existing antitrust doctrine was built around human conspiracies and lacks the conceptual vocabulary to address autonomous algorithmic convergence. The OECD recognised this gap in a landmark competition policy roundtable, warning that algorithms amplify the conditions for tacit coordination by increasing market transparency, enabling rapid retaliation, and reducing the benefits of deviation \citep{OECD2017}. This paper addresses the autonomous learning case, the fourth mode in the Ezrachi--Stucke taxonomy, which is both hardest to detect and most likely to grow in prevalence as algorithmic bidding becomes ubiquitous.

The central obstacle to enforcement is doctrinal. Both US antitrust law (Sherman Act \S1) and European competition law (Article~101 TFEU) require evidence of ``agreement'' or ``concerted practice'' to establish liability for horizontal coordination. When algorithms converge to supra-competitive outcomes through independent learning rather than explicit communication, this evidentiary threshold is not met. \citet{Harrington2018} proposes a structural remedy: a per se prohibition on pricing algorithms whose design incorporates features that facilitate collusion, shifting the legal focus from observed outcomes to algorithmic properties. \citet{Gal2019} argues that algorithms' inspectable and reproducible nature should expand rather than narrow the regulatory net, since algorithmic strategies can in principle be audited in ways that human decision-making cannot. \citet{Petit2017} offers a more sceptical view, contending that algorithmic heterogeneity in real markets may limit the scope for autonomous convergence, and that premature regulation risks chilling beneficial algorithmic innovation. \citet{Schwalbe2018} reinforces this caution, concluding that the theoretical conditions under which algorithmic collusion arises are more restrictive than the popular discussion suggests.

The enforcement landscape is evolving but has not yet addressed the autonomous case. In the United States, the Department of Justice's case against RealPage advances a hub-and-spoke theory of algorithmic coordination in rental housing markets. \citet{EzrachiStucke2024} generalise this concern through the concept of ``secondary algorithmic tacit collusion,'' showing that when multiple pricing hubs serve overlapping sets of competitors, conscious parallelism among the hubs can facilitate market-wide alignment even in differentiated product markets with many rivals. The Preventing Algorithmic Collusion Act has been introduced in Congress to create statutory authority over algorithm-facilitated price-fixing. In the European Union, the Court of Justice established in \emph{Eturas} (C-74/14) that participation in an information exchange through a shared algorithmic platform can constitute a concerted practice, and the 2023 Horizontal Cooperation Guidelines explicitly address algorithmic coordination for the first time. The UK Competition and Markets Authority has invested in algorithmic screening tools to detect collusive patterns in market data. Despite this activity, no jurisdiction has sanctioned purely autonomous tacit collusion by independent learning algorithms \citep{OECD2023, Harrington2018}. The gap between doctrinal requirements and algorithmic capabilities remains the central challenge.

The legal literature on algorithmic collusion focuses almost exclusively on pricing games, yet auctions present distinct challenges. Sealed-bid formats eliminate the price transparency that facilitates tacit coordination in posted-price markets, while intermediated platforms introduce a third party (the auctioneer or exchange) whose design choices shape competitive outcomes. Bid-rigging doctrine assumes explicit coordination among bidders, not autonomous convergence of learning algorithms. The finding by \citet{BanchioSkrzypacz2022} that first-price auctions are more vulnerable to algorithmic bid suppression reframes the policy debate: auction format itself becomes a regulatory instrument. \citet{Fish2024} extend algorithmic collusion to large language model agents, showing that LLM-based bidders reach supra-competitive outcomes in both pricing and auction settings, raising novel detection challenges because prompt-driven strategies are harder to audit than parametric algorithms. \citet{Heo2025} demonstrate that reinforcement learning agents can sustain collusive bidding patterns in simulated construction procurement auctions, extending the concern beyond digital advertising to public procurement.

Several reform strategies are relevant to auction markets. Algorithm-intrinsic approaches include \citeauthor{Harrington2018}'s per se prohibition and the calibrated regret auditing framework of \citet{Hartline2024, Hartline2025}, which proposes statistical tests to certify that a pricing algorithm's outcomes are consistent with competitive behaviour. Market design approaches treat the auction mechanism itself as a regulatory lever, using format choice, reserve prices, and information disclosure to constrain collusive convergence. Legislative approaches seek to expand the definition of concerted practice or create new statutory authority. At a broader level, \citet{Crane2024} argues that general-purpose AI threatens to undermine the four pillars on which the antitrust order is built, including information discovery, incentive alignment, market structure, and conduct control, suggesting that incremental doctrinal reform may prove insufficient as algorithmic capabilities continue to advance. Information-design approaches offer a complementary lever: \citet{Zhang2025Noise} proposes controlled noise injection into shared market data as an ex ante mechanism to disrupt algorithmic coordination without prohibiting the use of pricing algorithms themselves. The computational laboratory framework developed in this paper provides exactly the kind of systematic, controlled evidence that the legal literature identifies as missing: a quantitative ranking of which design parameters most effectively prevent learning-driven revenue shortfalls, grounded in orthogonal experimental variation rather than anecdotal case studies.

This paper contributes to all four literatures. First, prior algorithmic collusion work studies mechanisms in isolation; no systematic ranking of their relative importance exists. Second, auctions remain understudied relative to pricing games despite their economic significance. Third, the autobidding literature focuses on welfare bounds from the bidder's perspective rather than seller revenue outcomes. Fourth, and most critically, the policy-relevant finding that first-price auctions are more vulnerable to bid suppression rests entirely on experiments with algorithms that no major platform deploys; the present paper tests both Q-learning and pacing algorithm families under identical market conditions and shows that the auction-format effect reverses when the algorithm changes, a result that reframes the design recommendations that follow from the existing literature. The computational laboratory framework developed here addresses each of these gaps.
